Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-07-22 hybrid_placement.tex
tweak figures size

For the purpose of brevity, we only present the
communication time distribution of each application under all placement and routing configurations, including the hybrid placement method. These results are presented in Figure~\ref{fig:hyb-plcmt-apps-commtime}. 
As shown in Figure~\ref{fig:hyb-plcmt-mg-commtime} and~\ref{fig:hyb-plcmt-cr-commtime},
MultiGrid and CrystalRouter exhibit similar performance in both hybrid and random placement, as nodes are being placed randomly in each case.
While the performance of AMG under hybrid placement, shown in Figure~\ref{fig:hyb-plcmt-amg-commtime}, still appears to exhibit significant communication interference on account of the other applications as opposed to the best contiguous placement policies, the effects are significantly reduced compared to a random-adaptive policy. We believe this to be a result of more AMG-specific traffic occupying a smaller set of routers/groups, both reducing the probability of traffic entering them through adaptive routing and increasing the relative proportion of link utilization by AMG. Of course, this comes with the costs associated with contiguous allocation, in which AMG's traffic is less likely to load balance across multiple dragonfly groups.

>>>>

For the purpose of brevity, we only present the
communication time distribution of each application under all placement and routing configurations, including the hybrid placement method. These results are presented in Figure~\ref{fig:hyb-plcmt-apps-commtime}. 
As shown in Figure~\ref{fig:hyb-plcmt-mg-commtime} and~\ref{fig:hyb-plcmt-cr-commtime},
MultiGrid and CrystalRouter exhibit similar performance in both hybrid and random placement, as nodes are being placed randomly in each case.
While the performance of AMG under hybrid placement, shown in Figure~\ref{fig:hyb-plcmt-amg-commtime}, still appears to exhibit significant communication interference on account of the other applications as opposed to the best contiguous placement policies, the effects are reduced compared to a random-adaptive policy. We believe this to be a result of more AMG-specific traffic occupying a smaller set of routers/groups, both reducing the probability of traffic entering them through adaptive routing and increasing the relative proportion of link utilization by AMG. Of course, this comes with the costs associated with contiguous allocation, in which AMG's traffic is less likely to load balance across multiple dragonfly groups.

### 2016-07-22 study_workload_1.tex
tweak figures size

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/gc-traffic}
        \caption{Global Channel Traffic}
        \label{fig:global-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/lc-traffic}
        \caption{Local Channel Traffic}
        \label{fig:local-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/tl-traffic}
        \caption{Terminal Link Traffic}
        \label{fig:terminal-link-traffic}
    \end{subfigure}\\

    \centering   
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/gc-stime}
        \caption{Global Channel Saturation Time}
        \label{fig:global-channel-stime}
    \end{subfigure}\hfill
     \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/lc-stime}
        \caption{Local Channel Saturation Time}
        \label{fig:local-channel-stime}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/tl-stime}
        \caption{Terminal Link Saturation Time}
        \label{fig:terminal-link-stime}
    \end{subfigure}%
    \caption{Aggregate traffic and saturation time for Workload~\Rmnum{1} under the configurations listed in Table~\ref{tab: placement routing configs}. ``CA'' and ``CPA'' have equivalent behavior.}
   \label{fig:wkld-network-traffic-stime}
\end{figure*}

@@

The AMG application (Figure~\ref{fig:amg-commtime}), however, shows markedly
different behavior when using random allocation. Random allocation with minimal
routing results in worse performance than contiguous-adaptive configurations,
while using adaptive routing results in significant performance regressions. As
this is a counterintuitive result not discussed in other works, we investigate
further.

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/lc-traffic}
        \caption{MultiGrid Local Channel Traffic}
        \label{fig:mg-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/lc-traffic}
        \caption{CrystalRouter Local Channel Traffic}
        \label{fig:cr-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/lc-traffic}
        \caption{AMG Local Channel Traffic}
        \label{fig:amg-lc-traffic}
    \end{subfigure}\\  
    \centering
   \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/gc-traffic}
        \caption{MultiGrid Global Channel Traffic}
        \label{fig:mg-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/gc-traffic}
        \caption{CrystalRouter Global Channel Traffic}
        \label{fig:cr-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/gc-traffic}
        \caption{AMG Global Channel Traffic}
        \label{fig:amg-gc-traffic}
    \end{subfigure}
   \caption{Aggregate workload traffic for routers serving specific applications. ``CA'' and ``CPA'' have equivalent behavior. More routers are involved in serving each application when random placement is in use, compared to contiguous placement.}
   \label{fig:3app-lc-gc-traffic}
\end{figure*}

>>>>

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/gc-traffic}
        \caption{Global Channel Traffic}
        \label{fig:global-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/lc-traffic}
        \caption{Local Channel Traffic}
        \label{fig:local-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/tl-traffic}
        \caption{Terminal Link Traffic}
        \label{fig:terminal-link-traffic}
    \end{subfigure}\\

    \centering   
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/gc-stime}
        \caption{Global Channel Saturation Time}
        \label{fig:global-channel-stime}
    \end{subfigure}\hfill
     \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/lc-stime}
        \caption{Local Channel Saturation Time}
        \label{fig:local-channel-stime}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/tl-stime}
        \caption{Terminal Link Saturation Time}
        \label{fig:terminal-link-stime}
    \end{subfigure}%
    \caption{Aggregate traffic and saturation time for Workload~\Rmnum{1} under the configurations listed in Table~\ref{tab: placement routing configs}. ``CA'' and ``CPA'' have equivalent behavior.}
   \label{fig:wkld-network-traffic-stime}
\end{figure*}

@@

The AMG application (Figure~\ref{fig:amg-commtime}), however, shows markedly
different behavior when using random allocation. Random allocation with minimal
routing results in comparable performance to contiguous-adaptive configurations,
while using adaptive routing results in significant performance regressions. As
this is a counterintuitive result not discussed in other works, we investigate
further.

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/mg/lc-traffic}
        \caption{MultiGrid Local Channel Traffic}
        \label{fig:mg-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/cr/lc-traffic}
        \caption{CrystalRouter Local Channel Traffic}
        \label{fig:cr-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/amg/lc-traffic}
        \caption{AMG Local Channel Traffic}
        \label{fig:amg-lc-traffic}
    \end{subfigure}\\  
    \centering
   \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/mg/gc-traffic}
        \caption{MultiGrid Global Channel Traffic}
        \label{fig:mg-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/cr/gc-traffic}
        \caption{CrystalRouter Global Channel Traffic}
        \label{fig:cr-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/amg/gc-traffic}
        \caption{AMG Global Channel Traffic}
        \label{fig:amg-gc-traffic}
    \end{subfigure}
   \caption{Aggregate workload traffic for routers serving specific applications. ``CA'' and ``CPA'' have equivalent behavior. More routers are involved in serving each application when random placement is in use, compared to contiguous placement.}
   \label{fig:3app-lc-gc-traffic}
\end{figure*}

### 2016-07-22 study_workload_2.tex
tweak in Section 6
tweak figures size

In this section, we use a different experimental configuration to verify and
explore the observations made in the previous section. Specifically, we conduct
the same sets of experiments through Workload~\Rmnum{2}, which consists of sAMG,
MultiGrid and CrystalRouter. By replacing AMG with sAMG, we turn the ``bully"
into the ``bullied".

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/gc-traffic}
        \caption{Global Channel Traffic}
        \label{fig:synwkld-global-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/lc-traffic}
        \caption{Local Channel Traffic}
        \label{fig:synwkld-local-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/tl-traffic}
        \caption{Terminal Link Traffic}
        \label{fig:synwkld-terminal-link-traffic}
    \end{subfigure}\\

    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/gc-btime}
        \caption{Global Channel Saturation Time}
        \label{fig:synwkld-global-channel-stime}
    \end{subfigure}\hfill
     \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/lc-btime}
        \caption{Local Channel Saturation Time}
        \label{fig:synwkld-local-channel-stime}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{syn-wkld/tl-btime}
        \caption{Terminal Link Saturation Time}
        \label{fig:synwkld-terminal-link-stime}
    \end{subfigure}%
   \caption{Aggregate traffic and saturation time for Workload~\Rmnum{2} under the configurations listed in Table~\ref{tab: placement routing configs}.}
   \label{fig:synwkld-network-traffic-stime}
\end{figure*}

@@

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/mg/commtime}
        \caption{MultiGrid}
        \label{fig:syn-mg-commtime}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/cr/commtime}
        \caption{CrystalRouter}
        \label{fig:syn-cr-commtime}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/amg10/commtime}
        \caption{sAMG}
        \label{fig:syn-samg-commtime}
    \end{subfigure}%
    \caption{Communication time distribution across application ranks in Workload~\Rmnum{2}. The ``bully", sAMG, benefits from random placement and adaptive routing, while the ``bullied", MultiGrid and CrystalRouter, suffer performance degradation.}
   \label{fig:syn-apps-commtime}
\end{figure*}

@@

Interestingly, in both Workload~\Rmnum{1} and \Rmnum{2}, MultiGrid experiences
a more subtle performance variatioin than the significant swings in performance
observed in the other applications. These behaviors persisted across multiple runs
with different random seeds. Additionally, CrystalRouter in Workload~\Rmnum{2}
exhibits less drastic changes in maximal load, but still experiences performance
regressions. We are continuing to work towards understanding the root causes
and implications of this behavior, for which we expect application-specific
communication patterns to be an important factor. 
%\TODO{Is this a reasonable observation? Yes, it is.}

>>>>

In this section, we use a different experimental configuration to verify and
explore the observations made in the previous section. Specifically, we conduct
the same sets of experiments through Workload~\Rmnum{2}, which consists of sAMG,
MultiGrid and CrystalRouter.
%\NOTE{By replacing AMG with sAMG, we turn the ``bully" into the ``bullied".} 
The ``bully" becomes the ``bullied" with the presence of sAMG in Workload~\Rmnum{2}.

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/gc-traffic}
        \caption{Global Channel Traffic}
        \label{fig:synwkld-global-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/lc-traffic}
        \caption{Local Channel Traffic}
        \label{fig:synwkld-local-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/tl-traffic}
        \caption{Terminal Link Traffic}
        \label{fig:synwkld-terminal-link-traffic}
    \end{subfigure}\\

    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/gc-btime}
        \caption{Global Channel Saturation Time}
        \label{fig:synwkld-global-channel-stime}
    \end{subfigure}\hfill
     \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/lc-btime}
        \caption{Local Channel Saturation Time}
        \label{fig:synwkld-local-channel-stime}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/tl-btime}
        \caption{Terminal Link Saturation Time}
        \label{fig:synwkld-terminal-link-stime}
    \end{subfigure}%
   \caption{Aggregate traffic and saturation time for Workload~\Rmnum{2} under the configurations listed in Table~\ref{tab: placement routing configs}.}
   \label{fig:synwkld-network-traffic-stime}
\end{figure*}

@@

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/mg/commtime}
        \caption{MultiGrid}
        \label{fig:syn-mg-commtime}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/cr/commtime}
        \caption{CrystalRouter}
        \label{fig:syn-cr-commtime}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{syn-wkld/amg10/commtime}
        \caption{sAMG}
        \label{fig:syn-samg-commtime}
    \end{subfigure}%
    \caption{Communication time distribution across application ranks in Workload~\Rmnum{2}. The ``bully", sAMG, benefits from random placement and adaptive routing, while the ``bullied", MultiGrid and CrystalRouter, suffer performance degradation.}
   \label{fig:syn-apps-commtime}
\end{figure*}

@@

Interestingly, in both Workload~\Rmnum{1} and \Rmnum{2}, MultiGrid experiences
a more subtle performance variatioin than the significant swings in performance
observed in the other applications. These behaviors persisted across multiple runs
with different random seeds. Additionally, CrystalRouter in Workload~\Rmnum{2}
has less drastic changes in maximal load, but still experiences performance
regressions. We are continuing to work towards understanding the root causes
and implications of this behavior, for which we expect application-specific
communication patterns to be an important factor. 
%\TODO{Is this a reasonable observation? Yes, it is.}
